---
title: "Statistical Inference Course Project -  part 1"
date: "1 March 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

In this project you I’ll investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution is simulated in R with rexp(n, lambda) where lambda is the rate parameter equal 0.2. I will investigate the distribution of averages of 40 exponentials and there will be a thousand simulations.


```{r}
n <- 1000
samples <- 40
lambda <- 0.2

```

# Simulation

I will generate the simulation
```{r}
dataRaw <- matrix(data = rexp(n = n * samples, rate = lambda), nrow = n, ncol = samples)
dataRaw <- as.data.frame(dataRaw)
```

I will generate the distribution of the averages of 40 exponentials
```{r}
dataMean <- apply(X = dataRaw, MARGIN = 1, FUN = mean)
```

# Sample Mean versus Theoretical Mean and Sample Variance versus Theoretical Variance

### The theoretical mean of the distribution
```{r}
theoreticalMean <- 1 / lambda
theoreticalMean
```

### The sample mean
```{r}
sampleMean <- mean(dataMean)
sampleMean
```

The sample mean is close to the theoretical mean

### The theoretical variance of the distribution
```{r}
theoreticalVariance <- (1 / lambda)^2/samples
theoreticalVariance
```

### The sample variance
```{r}
sampleVariance <- var(dataMean)
sampleVariance
```

The sample variance is close to the theoretical variance


### The theoretical standard deviation of the distribution
```{r}
theoreticalSD <- 1/lambda/sqrt(samples)
theoreticalSD
```

### The sample standard deviation
```{r}
standardDeviation <- sd(dataMean)
standardDeviation
```

The sample standard deviation is close to the theoretical standard deviation

## The distribution of averages of 40 random eponentials with 1000 simulations

```{r include=FALSE}
library(ggplot2)
```

```{r}
ggplot(data = as.data.frame(dataMean), aes(dataMean)) +
  geom_histogram(aes(y=..density..), fill="white", col="lightblue", alpha=.3, bins = 30) +
  geom_density(col="lightblue3", lwd=2) +
  geom_vline(xintercept = sampleMean, col="lightblue2", linetype="dashed", lwd=2) +
  geom_vline(xintercept = theoreticalMean, col="red") +
  labs(x="Mean", y = "Density") +
  theme_light()
```

As you can see above the distribution of averages of independent and identically distributed (IID) variables becomes that of a standard normal.





